---
title: "Climate-Conflict Data Analysis"
author: "Boseong Yun and Jessica Anderson"
date: "4/25/2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries
library(tidymodels) 
library(lubridate) 
library(plm) # fixed effects package
library(fixest) # another fixed effects package with more intuitive standard errors
library(discrim) # Linear Discriminant Model
library(tidyverse)
library(here) # used for relative filepath
library(raster) # used for importing shapefile
library(ncdf4) # used for reading netCDF file
library(sf) # used for spatial join
```

# Importing subnational region data

```{r import-regions}
# Load in a shapefile of the regions of Cote d'Ivoire from the GADM website
# The commented code downloads the data from GADM and then imports it
#ci_regions <- raster::getData('GADM', country = "CIV", level = 2) %>%
#  st_as_sf(crs = 4326)
# The code below imports an RDS file already downloaded from GADM
ci_regions <- read_rds("gadm36_CIV_2_sp.rds") %>% 
  st_as_sf(crs = 4326) # using WGS84 coordinate reference system for all data

# Define new region id because existing GID_2 is not well formatted
# for future reference, appears that GID_2 was based on alphabetizing by NAME_1 
# and within that by NAME_2
ci_regions <- ci_regions %>% 
  mutate(
    region_id = str_replace_all(GID_2, "^CIV\\.", ""), # drop country code
    region_id = str_replace_all(region_id, "_1$", ""), # drop nonvarying characters
    region_id = str_replace_all(region_id, "\\.", ""), # delete decimal point
    region_id = as.integer(region_id) # convert from string to integer
    )
```

# Importing ACLED data on social disorder events

Not concerned about st_join and st_intersect_aw using planar coordinates because the distortion is likely to be minimal for the scale of our analysis and how close Cote d'Ivoire is to the equator.

```{r import-acled}
# Read the ACLED datafile of all events within Cote d'Ivoire
acled <- read_csv("ivory_coast.csv")

# Clean ACLED data and transform into spatial data format
acled_sf <- acled %>%
  # remove 48 events for which geo-precision may be broader than our regions
  filter(geo_precision != 3) %>%  
  # prepare year-month coding to match climate data
  mutate(
    date = dmy(event_date),
    month = month(date), 
    yrmo = floor_date(date, unit = "month")
  ) %>%
  # select variables of interest
  dplyr::select(
    year, month, yrmo, admin2, latitude, longitude, 
    event_type, sub_event_type, inter1, inter2, fatalities # inter1/2 = actor codes
    ) %>%
  # convert to simple features spatial data format
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# Add region IDs and names to ACLED data via spatial join
acled_regions_sf <- ci_regions %>%
  dplyr::select(geometry, NAME_2, region_id) %>% 
  st_join(acled_sf) # default left join will keep only events within Cote d'Ivoire

# Check accuracy of admin2 field
# it looks like GADM is more accurate in its region names than ACLED
# so we should collapse by region_id or NAME_2 when we aggregate
acled_regions_sf %>% dplyr::select(-geometry) %>% count(admin2, NAME_2) %>% view()

# Drop geometry because we no longer need it
acled_regions <- st_drop_geometry(acled_regions_sf) %>% as_tibble()

# Add indicator variables for event types and actors to facilitate aggregation
acled_regions <- acled_regions %>% 
  mutate(
    battles = if_else(event_type == "Battles", 1, 0),
    explosions = if_else(event_type == "Explosions/Remote violence", 1, 0),
    protests = if_else(event_type == "Protests", 1, 0),
    riots = if_else(event_type == "Riots", 1, 0),
    strategic = if_else(event_type == "Strategic developments", 1, 0), 
    violence_civ = if_else(event_type == "Violence against civilians", 1, 0), 
    state_forces = if_else(inter1 == 1 | inter2 == 1, 1, 0), 
    rebels = if_else(inter1 == 2 | inter2 == 2, 1, 0), 
    pol_militias = if_else(inter1 == 3 | inter2 == 3, 1, 0), 
    id_militias = if_else(inter1 == 4 | inter2 == 4, 1, 0), 
    rioters = if_else(inter1 == 5 | inter2 == 5, 1, 0), 
    protesters = if_else(inter1 == 6 | inter2 == 6, 1, 0), 
    external = if_else(inter1 == 8 | inter2 == 8, 1, 0)
  )

# Collapse into region-month panel
acled_panel <- acled_regions %>% 
  group_by(NAME_2, yrmo) %>% 
  summarise(
    region_id = mean(region_id), # keep region id in aggregated data
    events_count = n(), # count of all events (any type)
    battles_count = sum(battles),
    explosions_count = sum(explosions),
    protests_count = sum(protests),
    riots_count = sum(riots),
    strategic_count = sum(strategic), 
    violence_civ_count = sum(violence_civ), 
    state_forces_count = sum(state_forces), 
    rebels_count = sum(rebels), 
    pol_militias_count = sum(pol_militias), 
    id_militias_count = sum(id_militias), 
    rioters_count = sum(rioters), 
    protesters_count = sum(protesters), 
    external_count = sum(external),
    fatalities_count = sum(fatalities)
  ) %>% 
  ungroup()
```

# Importing ERA5 data from NetCDF file

```{r import-era5}
# Identify NetCDF filename
ncname <- "adaptor.mars.internal-1618978096.60298-4541-17-222d58bf-b1bc-4bf4-8fe0-1d49582df00e.nc"
# Import 2-meter temperature (t2m) from ERA5 (expver = 1)
temp <- brick(ncname, varname = "t2m", level = 1)
# Import total participation (tp) from ERA5 (expver = 1)
prcp <- brick(ncname, varname = "tp", level = 1)

# Define a function to convert each raster brick into tidy polygon geometry
raster_to_sf <- function(rasterbrick) {
  
  # Transform the raster brick into a polygon file
  polygons_sf <- rasterbrick %>%
    rasterToPolygons() %>% # min/max warnings come from this step
    st_as_sf(crs = 4326)
  
  # Save the name of the raster brick
  raster_name <- as.character(substitute(rasterbrick))
  
  # Tidy the polygon data
  clean_sf <- polygons_sf %>%
    # pivot months from wide to long
    # replacement length warnings come from this step
    pivot_longer( 
      -geometry,
      names_to = "date", 
      values_to = raster_name
      ) %>%
    # clean date format
    mutate(
      date = substr(date, 2, 11),
      date = ymd(date)
    ) %>%
    # drop missing values from last few months of data (related to ERA5 versions)
    drop_na() %>%
    st_as_sf(crs = 4326)
  
  # Return tidy polygon geometry
  return(clean_sf)
  
}

# Use function to tidy temperature and precipitation data
temp_sf <- raster_to_sf(temp)
prcp_sf <- raster_to_sf(prcp)
```

# Aggregating gridded climate data to subnational regions

```{r climate-interpolation}
# Set the dates parameter to loop over each year-month in the climate data
dates <- temp_sf %>% pull(date) %>% unique()

# Temperature ---------------------------------------------------------

# Create an output path for interpolated temperature data
output_temp <- list() 

# Loop over each year-month in the gridded temperature data
for (i in seq_along(dates)) {
  
  # filter the data to each year-month
  filtered <- temp_sf %>% filter(date == dates[i]) 
  
  # aggregate gridded temp to regions to get area-weighted average temp per region 
  interpolated <- st_interpolate_aw(
    filtered[, "temp"],
    ci_regions, 
    extensive = FALSE # average temp is spatially intensive (not a count)
    )
  
  # to the new temperature variable, add appropriate date and regions
  final <- interpolated %>% 
    mutate(
      date = ymd(dates[i]),
      region_name = ci_regions$NAME_2,
      region_id = ci_regions$region_id
    )
  
  # save each year-month of new temperature data to the output list
  output_temp[[i]] <- final
  
}

# Bind the months of interpolated temperature data into a long spatial dataframe
temp_interpolated <- bind_rows(output_temp)

# Precipitation ---------------------------------------------------------

# Create an output path for interpolated precipitation data
output_prcp <- list() 

# Loop over each year-month in the gridded precipitation data
for(i in seq_along(dates)) {
  
  # filter the data to each year-month
  filtered <- prcp_sf %>% filter(date == dates[i]) 
  
  # aggregate gridded prcp to regions to get area-weighted average prcp per region 
  interpolated <- st_interpolate_aw(
    filtered[, "prcp"],
    ci_regions, 
    extensive = FALSE) # average total precipitation is spatially intensive
  
  # to the new precipitation variable, add appropriate date and regions
  final <- interpolated %>% 
    mutate(
      date = ymd(dates[i]),
      region_name = ci_regions$NAME_2, 
      region_id = ci_regions$region_id
  )
  
  # save each year-month of new precipitation data to the output list
  output_prcp[[i]] <- final
  
}

# Bind the months of interpolated precipitation data into a long spatial dataframe
prcp_interpolated <- bind_rows(output_prcp)
```

# Merging interpolated climate data with conflict data by region

```{r merge-climate-and-conflict}
# Drop geometry from climate data and convert to more recognizable units
temp_clean <- temp_interpolated %>% 
  st_drop_geometry() %>% 
  as_tibble() %>% 
  rename(temp2m_k = temp) %>% 
  mutate(temp2m_c = temp2m_k - 273.15) # Kelvin to Celsius
prcp_clean <- prcp_interpolated %>% 
  st_drop_geometry() %>% 
  as_tibble() %>% 
  rename(prcp_m = prcp) %>% 
  mutate(prcp_mm = prcp_m * 1000) # meters to millimeters

# Merge conflict, temperature, and precipitation data on region and year-month
climate_conflict_merged <- temp_clean %>% 
  left_join(
    prcp_clean, 
    by = c("region_id", "region_name", "date")) %>% 
  left_join(
    acled_panel, 
    by = c("region_id" = "region_id", "region_name" = "NAME_2", "date" = "yrmo")
    )

# Replace missing ACLED counts with zeros and arrange data for readability
climate_conflict_region_month <- climate_conflict_merged %>% 
  dplyr::select(region_id, region_name, date, everything()) %>% 
  mutate(across(events_count:fatalities_count, ~replace_na(., 0))) %>% 
  arrange(region_id, region_name, date)

# Export full region-month panel dataset to CSV
write_csv(climate_conflict_region_month, "climate-conflict_region-month_panel.csv")
```

# Preparing data for modeling

```{r region-year}
# read in region-month panel data
region_month <- read_csv("climate-conflict_region-month_panel.csv")

# extract temperature and precipitation mean and sd from first 3 years of data
# in the future we'll use climate data from 1987 to 1996
climate_baselines <- region_month %>% 
  mutate(year = year(date)) %>% 
  filter(dplyr::between(year, 1997, 1999)) %>% 
  group_by(region_id) %>% 
  summarise(
    mean_temp2mc_baseline = mean(temp2m_c), 
    mean_prcpmm_baseline = mean(prcp_mm), 
    sd_temp2mc_baseline = sd(temp2m_c),
    sd_prcpmm_baseline = sd(prcp_mm)
  )

# create standardized climate variables (standard deviational units from 1997 mean)
region_month <- region_month %>% 
  left_join(climate_baselines, by = "region_id") %>% 
  mutate(
    temp2m_c_std = (temp2m_c - mean_temp2mc_baseline) / sd_temp2mc_baseline, 
    prcp_mm_std = (prcp_mm - mean_prcpmm_baseline) / sd_prcpmm_baseline
  )

# collapse to region-year panel
region_year <- region_month %>% 
  mutate(year = year(date)) %>% 
  group_by(region_name, year) %>% 
  summarise(
    region_id = mean(region_id), # keep region id in aggregated data
    temp2m_c = mean(temp2m_c),
    prcp_mm = mean(prcp_mm),
    temp2m_c_std = mean(temp2m_c_std),
    prcp_mm_std = mean(prcp_mm_std),
    events_count = sum(events_count), # count of all events (any type)
    battles_count = sum(battles_count),
    explosions_count = sum(explosions_count),
    protests_count = sum(protests_count),
    riots_count = sum(riots_count),
    strategic_count = sum(strategic_count), 
    violence_civ_count = sum(violence_civ_count), 
    state_forces_count = sum(state_forces_count), 
    rebels_count = sum(rebels_count), 
    pol_militias_count = sum(pol_militias_count), 
    id_militias_count = sum(id_militias_count), 
    rioters_count = sum(rioters_count), 
    protesters_count = sum(protesters_count), 
    external_count = sum(external_count),
    fatalities_count = sum(fatalities_count)
  ) %>% 
  ungroup() %>% 
  dplyr::select(region_id, everything())
```

# Initial models

```{r plm-models}
# separate models with unit FE
temp_unit <- plm(
  formula = events_count ~ temp2m_c_std,
  data = region_year,
  effect = "individual", 
  model = "within",
  index = "region_id"
)
prcp_unit <- plm(
  formula = events_count ~ prcp_mm_std,
  data = region_year,
  effect = "individual", 
  model = "within", 
  index = "region_id"
)

# combined model with unit FE
both_unit <- plm(
  formula = events_count ~ temp2m_c_std + prcp_mm_std,
  data = region_year,
  effect = "individual", 
  model = "within", 
  index = "region_id"
)

# combined model with unit and time FE
both_twoway <- plm(
  formula = events_count ~ temp2m_c_std + prcp_mm_std,
  data = region_year,
  effect = "twoways", 
  model = "within", 
  index = c("region_id", "year")
)

# model summaries
summary(temp_unit)
summary(prcp_unit)
summary(both_unit)
summary(both_twoway)
```

```{r fixest-models}
temp_unit_feols <- feols(
  events_count ~ temp2m_c_std | region_id, # by default clusters SEs on region_id
  data = region_year
)
prcp_unit_feols <- feols(
  events_count ~ prcp_mm_std | region_id, # by default clusters SEs on region_id
  data = region_year
)
both_unit_feols <- feols(
  events_count ~ temp2m_c_std + prcp_mm_std | region_id, # clusters on region_id
  data = region_year
)
both_twoway_feols <- feols(
  events_count ~ temp2m_c_std + prcp_mm_std | region_id + year,
  data = region_year,
  cluster = ~region_id + year
)

summary(temp_unit_feols)
summary(prcp_unit_feols)
summary(both_unit_feols)
summary(both_twoway_feols)
```
